---
title: "p8105_hw2_zz3039"
author: "Zirui Zhang"
date: "2022-10-02"
output: github_document
---
```{r chunk_library, message=FALSE}
library(tidyverse)
library(tidyr)
library(readr)
library(haven)
library(readxl)
options(tibble.print_min = 5)
```

## Problem 2
```{r chunk_Problem2_mr}
wheel_data1 =
  read_excel(
    "./data/Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    skip = 1,
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  rename(
    weight = weight_tons,
    volume = volume_cubic_yards,
    PlasticBottles = plastic_bottles,
    Polystyrene = polystyrene,
    CigaretteButts = cigarette_butts,
    GlassBottles = glass_bottles,
    GroceryBags = grocery_bags,
    ChipBags = chip_bags,
    SportsBalls = sports_balls,
    HomesPowered = homes_powered) %>% 
  drop_na(dumpster) %>% 
  mutate(WheelName = "Mr. Trash Wheel",
         SportsBalls =round(SportsBalls, digits = 0),
         SportsBalls = as.integer(SportsBalls)) %>% 
  select(dumpster, date, year, month, WheelName, everything())
wheel_data1
```

```{r chunk_Problem2_professor}
wheel_data2 =
  read_excel(
    "./data/Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    skip = 1,
    range = cell_cols("A:M")) %>% 
  janitor::clean_names() %>% 
  rename(
    weight = weight_tons,
    volume = volume_cubic_yards,
    PlasticBottles = plastic_bottles,
    Polystyrene = polystyrene,
    CigaretteButts = cigarette_butts,
    GlassBottles = glass_bottles,
    GroceryBags = grocery_bags,
    ChipBags = chip_bags,
    HomesPowered = homes_powered) %>% 
  drop_na(dumpster) %>% 
  mutate(
    WheelName = "Professor Trash Wheel",
    year = as.character(year)) %>%
  select(dumpster, date, year, month, WheelName, everything())
wheel_data2
```

```{r chunk_Problem2_join}
wheel_bind = 
  bind_rows(wheel_data1, wheel_data2) 
wheel_bind
```
* The total number of observations in this data is **`r nrow(wheel_bind)`**, the key variables include: ***`r names(wheel_bind)`***;

* The total weight of trash collected by Professor Trash Wheel is **`r sum(wheel_bind$weight[wheel_bind$WheelName == 'Professor Trash Wheel'])`** tons.

* The total number of sports balls collected by Mr. Trash Wheel in 2020 was **`r sum(wheel_bind$SportsBalls[wheel_bind$year == 2020], na.rm = TRUE)`**.


## Problem 3

```{r chunk_pols, warning=FALSE, message=FALSE}
pols_month_data = read_csv("./data/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.integer(year),
         day = as.integer(day),
         month = as.integer(month),
         month = month.abb[month],
         month = factor(month, levels = month.abb)) %>% 
  mutate(prez = prez_gop - prez_dem) %>%
  mutate(
    president = recode(
      prez,
      "1" = "republican",
      "-1" = "democratic"
    )
  ) %>% 
  select(- prez, -prez_gop, - prez_dem, -day) %>%
  arrange(year, month) %>% 
  select(year, month, everything())
pols_month_data
```

```{r chunk_snp, warning=FALSE, message=FALSE}
snp_data = read_csv("./data/snp.csv") %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  select(-day) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         month = month.abb[month], 
         month = factor(month, levels = month.abb)) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(
    year = 
      case_when(
        year <= 22 ~ year + 2000,
        year > 22 ~ year + 1900)) %>%
  arrange(year, month) %>% 
  select(year, month, everything())
snp_data
```

```{r chunk_unem, warning=FALSE, message=FALSE}
unem_data = read_csv("./data/unemployment.csv") %>%
  pivot_longer(
      Jan:Dec,
      names_to = "month",
      values_to = "prct"
    ) %>% 
  janitor::clean_names() %>% 
  mutate(month = factor(month, levels = month.abb)) %>% 
  arrange(year, month) %>% 
  select(year, month, everything())
unem_data
```

```{r chunk_join}
join1 =
  merge(
    x = pols_month_data,
    y = snp_data,
    by.x = c("year", "month"),
    by.y = c("year", "month"))
join2 = 
  merge(
    x = join1,
    y = unem_data,
    by.x = c("year", "month"),
    by.y = c("year", "month")
  )
```

* The **_pols_month_data_** dataset contained the number of ***governors***, ***senators***, ***representatives*** of the two parties, while also showing the ***party of the president*** on the associate date.

* The **_snp_data_** dataset contained the ***Standard & Poor’s stock market index*** on the associated date.

* The **_unem_data_** contained the ***percentage of unemployment*** on the associate date.

* The resulting dataset **_join2_** is a combination of the **3** datasets above, it has **`r ncol(join2)`** dimensions with a year range across **`r min(join2$year)`** to **`r max(join2$year)`**. The key variables in **_join2_** include ***`r names(join2)`***.

